---
title: "Marketing Insight Report - Facebook Metrics"
author: "Etini Akpayang"
date: "2025-01-30"
output: html_document
---
## Loading Libraries

```{r setup, include=FALSE}
library(tidyverse)
library(ggplot2)
```

# Introduction

We will briefly attempt to get key marketable metrics from this the facebook data to aid the marketing department develop data driven strategies going forward.


## Data Loading, Viewing, and Structure Understanding

```{r}
fb_data <- read.csv2("~/CODE/ANALYSIS/pde/Etini stuff/hng12 stuff/dataset_Facebook.csv")
glimpse(fb_data)
head(fb_data, 3)
```

## Data Cleaning

We would check the data for consistency and credibility

```{r}
# Clean column names
colnames(fb_data) <- gsub("\\.", "_", colnames(fb_data))

# Convert categorical variables
fb_data <- fb_data %>%
  mutate(
    Paid = factor(Paid, levels = c(0, 1), labels = c("Organic", "Paid")),
    Post_Month = factor(Post_Month, levels = 1:12, labels = month.abb),
    Post_Weekday = factor(Post_Weekday, levels = 1:7, labels = c("Sun", "Mon", "Tue", "Wed", "Thu", "Fri", "Sat"))
  ) %>% 
  drop_na()

glimpse(fb_data)

# Missing data check
data.frame(
  Variable = colnames(fb_data),
  Missing_Values = colSums(is.na(fb_data))
)
```

## Key Engagement Insights
How is this pages activity metric? is it marketable?
```{r}
engagement_summary <- fb_data %>% 
  summarise(
    avg_like = mean(like, na.rm = TRUE),
    avg_comments = mean(comment, na.rm = TRUE),
    avg_shares = mean(share, na.rm = TRUE)
  )
engagement_summary
```

## Top Performing Posts
What are the toptop performing post on the page?
```{r}
top_posts <- fb_data %>% 
  arrange(desc(like)) %>% 
  select(Type, like, comment, share, Post_Hour, Post_Month, Post_Weekday) %>% 
  head(10)
top_posts
```

## Engagement by Post Type
What does the engagemnts on our posts look like?

```{r}
fb_data %>%
  group_by(Type) %>%
  summarise(Avg_Interactions = mean(Total_Interactions, na.rm = TRUE)) %>%
  ggplot(aes(x = reorder(Type, Avg_Interactions), y = Avg_Interactions, fill = Type)) +
  geom_col(show.legend = FALSE) +
  coord_flip() +
  labs(title = "Engagement by Post Type", x = "Post Type", y = "Average Interactions") +
  theme_minimal()
```

## Best Posting Time for Engagement
Could there be a significantly better time to make a post?

```{r}
fb_data %>%
  group_by(Post_Hour) %>%
  summarise(Avg_Interactions = mean(Total_Interactions, na.rm = TRUE)) %>%
  ggplot(aes(x = Post_Hour, y = Avg_Interactions)) +
  geom_line(color = "blue", size = 1) +
  geom_point(color = "red") +
  labs(title = "Engagement by Posting Hour", x = "Hour", y = "Average Interactions") +
  theme_minimal()
```

## Paid vs. Organic Performance

lets consider that cisuals on paid posts and organic posts is there something to uncover
```{r}
fb_data %>%
  group_by(Paid) %>%
  summarise(Avg_Reach = mean(Lifetime_Post_Total_Reach, na.rm = TRUE),
            Avg_Interactions = mean(Total_Interactions, na.rm = TRUE)) %>%
  pivot_longer(cols = c(Avg_Reach, Avg_Interactions), names_to = "Metric", values_to = "Value") %>%
  ggplot(aes(x = Paid, y = Value, fill = Paid)) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~ Metric, scales = "free") +
  theme_minimal() +
  labs(title = "Paid vs. Organic Post Performance", x = "Paid Status", y = "Metric Value")
```

## Conclusion

This report highlights key insights into Facebook post performance, including **top-performing posts**, **engagement trends**, and comparisons between **paid vs. organic** strategies. It highlights that the best posts that had most engagements were videos and paid post always always outperformed organic post in interaction and engagements. Video contents should be encouraged and links can be included in them as when posted alone have significantly low interactions. and finally the magic time of the 5th hour should be exploited.

## Recommendations

More analytics can be don to know who exactly to target to get optimal interactuion and engagements and the days and months included could have deeper meaning is critically accessed.

---
